Taking Dyads Seriously

### 2018-02-12 paper/4_app.tex
app section

\section{\textbf{Re-estimation with AME}}

@@

In order to showcase the insights the AME approach permits scholars to bring into their analysis, we apply AME to five recent studies from the broad field of international relations and international political economy~\citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}. Each of these studies are representative of broader trends in the field in that they use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system.

We chose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the AME approach. Firstly, the results of AME estimation are interpretable alongside GLM results but have the additional benefit of being able to quantify higher-order features of interest to scholars, such as ``sender'' and ``receiver'' effects. Secondly, although the data requirements of AME are, admittedly, more intensive than needed for GLMs, ever richer data is increasingly prevalent in both new and existing data. Thirdly, through AME, we can examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.

The five studies replicated below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline. Of course, by the criteria of influence many other papers could have been considered for the replication. However, as described in the following sections, the AME model has specific data requirements, which limited the scope of potential papers.}  Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential. Reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized regression.

In the following section, we present two sets of results for each study, one that replicates the original study and another that tests the central hypothesis within an AME implementation. We demonstrate that AME allows scholars the ability to test their hypotheses of interest, while directly accounting for both theorized and observed interdependencies in the data. As accounting for interdependences can change the outcome of regression models, both in substance and in estimate, we also summarize the predictive performance of the GLM and AME implementations. This predictive view allows us to better compare the models across specifications (and where they differ in their coefficients, see which specification does better at modeling the underlying data-generating process). Finally, for several of the studies, we report the AME-derived sender and receiver effects in order to highlight additional value of the AME framework.

@@

% SM: maybe too on the nose (Whether putative control variables, for
% example, are precisely estimated is less important.) 
%% MJF: 2/9-- took this part out, because I don't think that it's necessar.
% Typically, scholars are concerned with conducting hypothesis testing on a small set of independent variables that they have assembled in a complicated empirical model. For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of a dispute dyad being in a dispute.  In a regression including fourteen covariates, they focus on whether the democratic nature of the initiating country is a statistically significant covariate.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framewo
%rk.

@@

% as a result of the incompatibility between GLM models and the underlying processes that shape international events, 
%MJF: Keeping this here for now, but it reads more like a conclusion

@@

\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes. Their work contests prior scholarship that had claimed interstate dyads containing democracies and personalist dictatorships were particularly prone to conflict because of aggression on the part of the democratic state. Using directed dyads, they find evidence against this hypothesis: dictators are in fact more likely to challenge democracies, but not the other way around.  In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, but the opposite is not true. Independent variables are largely taken from an earlier study and focus on various encodings of regime types, contiguity, alliance, and capability measures. As is prevalent in this literature, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines. The database for this study is constructed using \texttt{EUGene} \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads. Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.

>>>>

\section{\textbf{Applications with AME}}

@@

In order to showcase the insights the AME approach permits scholars to bring into their analysis, we apply AME to five recent studies from international relations and international political economy: \citet{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}. Each of these studies are representative of broader trends in the field in that they use relational data of state interactions and propose both dyadic, monadic, and structural explanations for behavior of actors in the system. We chose to demonstrate the capabilities of AME with reference to existing studies in order to highlight several features of the AME approach. First, the results of AME estimation are interpretable alongside GLM results but have the additional benefit of being able to quantify higher-order features of interest to scholars, such as ``sender'' and ``receiver'' effects. Second, through AME, we can examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence. Third, we will show that by using the AME framework scholars can better model the data generating process behind their events of interest.

The five studies used below were selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline. Of course, by the criteria of influence many other papers could have been considered for the replication. However, as described in the following sections, the AME model has specific data requirements, which limited the scope of potential papers.}  Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential. Reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized model.

% In the following section, we present two sets of results for each study, one that replicates the original study and another that tests the central hypothesis within an AME implementation. We demonstrate that AME allows scholars the ability to test their hypotheses of interest, while directly accounting for both theorized and observed interdependencies in the data. As accounting for interdependences can change the outcome of regression models, both in substance and in estimate, we also summarize the predictive performance of the GLM and AME implementations. This predictive view allows us to better compare the models across specifications (and where they differ in their coefficients, see which specification does better at modeling the underlying data-generating process). Finally, for several of the studies, we report the AME-derived sender and receiver effects in order to highlight additional value of the AME framework.

@@

@@

@@

\citet{reiter:stam:2003} examine the relationship between democracy, dictatorship and the initiation of militarized disputes. Their work contests prior scholarship that had claimed interstate dyads containing democracies and personalist dictatorships were particularly prone to conflict because of aggression on the part of the democratic state. Using directed dyads, they find evidence against this hypothesis: dictators are in fact more likely to challenge democracies, but not the other way around.  In addition, military regimes and single-party regimes are more prone to initiate disputes with democracies, but the opposite is not true. Independent variables focus on various encodings of regime types, contiguity, alliance, and capability measures. As is prevalent in this literature, Reiter \& Stam employ a logistic regression that includes an indicator of the time since the last dispute as well as three cubic splines. The database for this study is constructed using \texttt{EUGene} \citep{bennett:stam:2000} and comprises approximately three-quarters of a million stacked dyads. Based on their statistical analysis, they conclude that institutional constraints affect the propensity of democratic and non-democratic leaders to engage in military conflict.
